---
title: "Chapter 3 - SQL Queries"
author: "meelyn"
date: "2025-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = FALSE)
```

# 3 SQL Queries

To obtain the data in your database, you will need to use SQL queries to get data from different tables.

This chapter is a quick summary on how to use SQL in R. If you would like more info on how to run different queries, use this tutorial: [https://solutions.posit.co/connections/db/getting-started/database-queries/](https://solutions.posit.co/connections/db/getting-started/database-queries/) 

## 3.1 List Tables

List the tables in your database. If everything was run correctly, each data file type (raw, blu, node-health, etc.) should be in its own data table.

Remember to reconnect to the database!

```{r}
library(celltracktech)

con <- DBI::dbConnect(duckdb::duckdb(), 
                      dbdir = "./data/meadows/meadows2.duckdb", 
                      read_only = FALSE)

# list tables in database
DBI::dbListTables(con)

# disconnect from database
DBI::dbDisconnect(con)

```
## 3.2 Load each data table into RStudio

```{r}
# connect to database
con <- DBI::dbConnect(duckdb::duckdb(), 
                      dbdir = "./data/meadows/meadows.duckdb", 
                      read_only = FALSE)

# list first 10 records in raw
raw = DBI::dbGetQuery(con, "SELECT * FROM raw ")
head(raw)

# list first 10 records in blu
blu = DBI::dbGetQuery(con, "SELECT * FROM blu ")
head(blu)

#list first 10 records in gps table
gps = DBI::dbGetQuery(con, 'SELECT * FROM gps')
head(gps)

# list first 10 records in node_health table
node_health = DBI::dbGetQuery(con, 'SELECT * FROM node_health')

# list the number of unique nodes in your project
unique_nodes = DBI::dbGetQuery(con, 'SELECT * FROM nodes')
head(unique_nodes)

# list the data files that were used to create your database
df_table = DBI::dbGetQuery(con, 'SELECT * FROM data_file')

# disconnect from database
DBI::dbDisconnect(con)

```

Note: you do not need to load each table into RStudio. You should only load the tables you need to for your analysis.